---
title: "wykresy"
output:
  html_document:
    df_print: paged
---



```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

```



```{r warning=F, include=F}
library(tidyverse)
library(ggforce)
library(ggridges)
library(ggthemes)
library(viridis)

```


```{r include=F}
# df_raw <- openxlsx::read.xlsx('../../Dane/Dane-08-06-2022a.xlsx') %>%
#   janitor::clean_names()

df_raw <- openxlsx::read.xlsx('../../Dane/2023-09-07 hormony.xlsx') %>%
  janitor::clean_names()



df <- df_raw %>% 
  mutate(grupa = case_when(str_detect(id_osoby, "B") ~ "EG1",
                           str_detect(id_osoby, "Z") ~ "EG2",
                           TRUE ~ "CG"),
         .before = id_osoby) %>%  
  mutate(
    id_pomiaru = case_when(
      id_pomiaru == 1 ~ 1,
      id_pomiaru == 2 ~ 6,
      id_pomiaru == 3 ~ 12,
      id_pomiaru == 4 ~ 16,
    ),
    grupa = factor(grupa, levels = c("EG1", "EG2", "CG")),
    id_osoby = factor(id_osoby),
    id_pomiaru = factor(id_pomiaru),
    nazwisko = factor(nazwisko)
  ) 


 # df %>% 
 #   group_by(grupa, id_pomiaru) %>% 
 #   summarize(n = n())

# skimr::skim(df)

# df %>%
# colnames()
```



```{r include=F}
df %>% 
  colnames()
```

# Nowe

Dane
```{r}


grupy = c("EG1","EG2", "CG")

var = "leptyna_ng_lml"
title = "Leptyna"
title_to_save = "leptyna"

x_title = "tydzień badań"
y_title = "[ng/ml]"


```



```{r}


plot <- df %>% 
  filter(
    grupa %in% grupy,
    .data[[var]] > 0
    ) %>% 
  # select(.data[[var]])
  ggplot(aes(x = id_pomiaru, y = .data[[var]], color = grupa)) +
  geom_boxplot(alpha = 0.7, 
            show.legend = F) +
  # stat_summary(fun.y=mean, geom="point", size=3, color = "red", aes(shape = "średnia"))+
  # scale_shape_manual(values = c("średnia"=20))+  
  geom_sina(aes(color = grupa),
            alpha = 0.25,
            show.legend = F)+
  facet_wrap(~grupa)+
  theme_classic() +
  theme(plot.title = element_text(face = "bold.italic", hjust = 0.5),
        plot.subtitle = element_text(face = 'italic', hjust = 0.5),
        legend.position = 'right',
        panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white")) +
  labs(title = title,
       x = x_title,
       y = y_title)+
  # scale_y_continuous(limits = c(NA, 37))+
  scale_color_tableau("Tableau 10")



plot
```

```{r}
ggsave(paste0(title_to_save, "_box_.png"), width = 8, height = 6, plot = plot, dpi = 300)

```

Tableau 10, Tableau 20, Color Blind, Seattle Grays, Traffic, Miller Stone, Superfishel Stone, Nuriel Stone, Jewel Bright, Summer, Winter, Green-Orange-Teal, Red-Blue-Brown, Purple-Pink-Gray, Hue Circle, Classic 10, Classic 10 Medium, Classic 10 Light, Classic 20, Classic Gray 5, Classic Color Blind, Classic Traffic Light, Classic Purple-Gray 6, Classic Purple-Gray 12, Classic Green-Orange 6, Classic Green-Orange 12, Classic Blue-Red 6, Classic Blue-Red 12, Classic Cyclic.



test nowego wykresu (że średnimi i odchyleniami st zamiast mediany i Q75)

```{r include=F}
df %>% 
  colnames()
```

Dane
```{r}


grupy = c("EG1","EG2", "CG")

var = "il_8_pg_ml"
title = "Il 8"
title_to_save = "il_8"

x_title = "tydzień badań"
y_title = "[pg/ml]"


```


```{r}
# definicja funkcji pomocniczej
min_mean_sd_max <- function(x) {
  r <- c(min(x), mean(x) - sd(x), mean(x), mean(x) + sd(x), max(x))
  names(r) <- c("ymin", "lower", "middle", "upper", "ymax")
  r
}


p1 <- df %>% 
  filter(
    grupa %in% grupy,
    .data[[var]] > 0
    ) %>% 
  ggplot(aes(x = id_pomiaru, y = .data[[var]], color = grupa))

p2 <- p1 +
  stat_summary(fun.data = min_mean_sd_max, geom = "boxplot", show.legend = F) +
  geom_sina(aes(color = grupa),
            alpha = 0.25,
            show.legend = F)+
  facet_wrap(~grupa)+
  theme_classic() +
  theme(plot.title = element_text(face = "bold.italic", hjust = 0.5),
        plot.subtitle = element_text(face = 'italic', hjust = 0.5),
        legend.position = 'right',
        panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white")) +
  labs(title = title,
       x = x_title,
       y = y_title)+
  scale_color_tableau("Tableau 10")
  
  
p2
```

```{r}
ggsave(paste0(title_to_save, "_box_mean_sd_v2.png"), width = 8, height = 6, plot = p2, dpi = 300)
```



Poniższy wykres do obczajenia:

```{r}
ggplot_box_legend <- function(family = "serif"){

  # Create data to use in the boxplot legend:
  set.seed(100)

  sample_df <- data.frame(parameter = "test",
                        values = sample(500))

  # Extend the top whisker a bit:
  sample_df$values[1:100] <- 701:800
  # Make sure there's only 1 lower outlier:
  sample_df$values[1] <- -350

  # Function to calculate important values:
  ggplot2_boxplot <- function(x){

    quartiles <- as.numeric(quantile(x,
                                     probs = c(0.25, 0.5, 0.75)))

    names(quartiles) <- c("25th percentile",
                          "50th percentile\n(median)",
                          "75th percentile")

    IQR <- diff(quartiles[c(1,3)])

    upper_whisker <- max(x[x < (quartiles[3] + 1.5 * IQR)])
    lower_whisker <- min(x[x > (quartiles[1] - 1.5 * IQR)])

    upper_dots <- x[x > (quartiles[3] + 1.5*IQR)]
    lower_dots <- x[x < (quartiles[1] - 1.5*IQR)]

    return(list("quartiles" = quartiles,
                "25th percentile" = as.numeric(quartiles[1]),
                "50th percentile\n(median)" = as.numeric(quartiles[2]),
                "75th percentile" = as.numeric(quartiles[3]),
                "IQR" = IQR,
                "upper_whisker" = upper_whisker,
                "lower_whisker" = lower_whisker,
                "upper_dots" = upper_dots,
                "lower_dots" = lower_dots))
  }

  # Get those values:
  ggplot_output <- ggplot2_boxplot(sample_df$values)

  # Lots of text in the legend, make it smaller and consistent font:
  update_geom_defaults("text",
                     list(size = 3,
                          hjust = 0,
                          family = family))
  # Labels don't inherit text:
  update_geom_defaults("label",
                     list(size = 3,
                          hjust = 0,
                          family = family))

  # Create the legend:
  # The main elements of the plot (the boxplot, error bars, and count)
  # are the easy part.
  # The text describing each of those takes a lot of fiddling to
  # get the location and style just right:
  explain_plot <- ggplot() +
    stat_boxplot(data = sample_df,
                 aes(x = parameter, y=values),
                 geom ='errorbar', width = 0.3) +
    geom_boxplot(data = sample_df,
                 aes(x = parameter, y=values),
                 width = 0.3, fill = "lightgrey") +
    geom_text(aes(x = 1, y = 950, label = "500"), hjust = 0.5) +
    geom_text(aes(x = 1.17, y = 950,
                  label = "Number of values"),
              fontface = "bold", vjust = 0.4) +
    theme_minimal(base_size = 5, base_family = family) +
    geom_segment(aes(x = 2.3, xend = 2.3,
                     y = ggplot_output[["25th percentile"]],
                     yend = ggplot_output[["75th percentile"]])) +
    geom_segment(aes(x = 1.2, xend = 2.3,
                     y = ggplot_output[["25th percentile"]],
                     yend = ggplot_output[["25th percentile"]])) +
    geom_segment(aes(x = 1.2, xend = 2.3,
                     y = ggplot_output[["75th percentile"]],
                     yend = ggplot_output[["75th percentile"]])) +
    geom_text(aes(x = 2.4, y = ggplot_output[["50th percentile\n(median)"]]),
              label = "Interquartile\nrange", fontface = "bold",
              vjust = 0.4) +
    geom_text(aes(x = c(1.17,1.17),
                  y = c(ggplot_output[["upper_whisker"]],
                        ggplot_output[["lower_whisker"]]),
                  label = c("Largest value within 1.5 times\ninterquartile range above\n75th percentile",
                            "Smallest value within 1.5 times\ninterquartile range below\n25th percentile")),
                  fontface = "bold", vjust = 0.9) +
    geom_text(aes(x = c(1.17),
                  y =  ggplot_output[["lower_dots"]],
                  label = "Outside value"),
              vjust = 0.5, fontface = "bold") +
    geom_text(aes(x = c(1.9),
                  y =  ggplot_output[["lower_dots"]],
                  label = "-Value is >1.5 times and"),
              vjust = 0.5) +
    geom_text(aes(x = 1.17,
                  y = ggplot_output[["lower_dots"]],
                  label = "<3 times the interquartile range\nbeyond either end of the box"),
              vjust = 1.5) +
    geom_label(aes(x = 1.17, y = ggplot_output[["quartiles"]],
                  label = names(ggplot_output[["quartiles"]])),
              vjust = c(0.4,0.85,0.4),
              fill = "white", label.size = 0) +
    ylab("") + xlab("") +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid = element_blank(),
          aspect.ratio = 4/3,
          plot.title = element_text(hjust = 0.5, size = 10)) +
    coord_cartesian(xlim = c(1.4,3.1), ylim = c(-600, 900)) +
    labs(title = "EXPLANATION")

  return(explain_plot)

}

plot <- ggplot_box_legend()

plot
```

```{r}
ggsave("legenda.png", plot, height = 8, width = 10, dpi = 300)
```



```{r}
for (i in 1:4){
  
  
  pomiar = c(i)
  subtitle = paste("pomiar", pomiar)
    
  plot <- df %>% 
    filter(
      id_pomiaru == pomiar,
      grupa %in% grupy) %>% 
  ggplot(aes(x = grupa, y = .data[[var]], color = grupa)) +
  geom_boxplot(alpha = 0.5, 
            show.legend = F) +
  stat_summary(fun.y=mean, geom="point", size=3, color = "red", aes(shape = "średnia"))+
  scale_shape_manual(values = c("średnia"=20))+  
  geom_sina(aes(color = grupa),
            alpha = 0.25,
            show.legend = F)+
  theme_classic() +
  theme(plot.title = element_text(face = "bold.italic", hjust = 0.5),
        plot.subtitle = element_text(face = 'italic', hjust = 0.5),
        legend.position = 'right',
        panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white")) +
  labs(title = title,
       x = x_title,
       y = y_title,
       shape="")+
  guides(color = guide_legend(override.aes = list(shape = c("średnia"=16))))  +
  scale_color_tableau()
  
  ggsave(paste0(title_to_save, "_box.png"), width = 8, height = 6, plot = plot, dpi = 300)
  
  
}
```






# Stare

# wykresy pomiędzy grupami

Dane
```{r}
var = "irisyna_ng_ml"

grupy = c("badana 1", "badana 2", "kontrolna")

title = "Iryzyna"
title_to_save = "iryzyna"

x_title = "Grupy"
y_title = "ng/ml"

i = 1
```



```{r}
for (i in 1:4){
  
  
  pomiar = c(i)
  subtitle = paste("pomiar", pomiar)
    
  plot <- df %>% 
    filter(
      id_pomiaru == pomiar,
      grupa %in% grupy) %>% 
  ggplot(aes(x = grupa, y = .data[[var]], color = grupa)) +
  geom_boxplot(alpha = 0.5, 
            show.legend = F) +
  stat_summary(fun.y=mean, geom="point", size=3, color = "red", aes(shape = "średnia"))+
  scale_shape_manual(values = c("średnia"=20))+  
  geom_sina(aes(color = grupa),
            alpha = 0.25,
            show.legend = F)+
  theme_classic() +
  theme(plot.title = element_text(face = "bold.italic", hjust = 0.5),
        plot.subtitle = element_text(face = 'italic', hjust = 0.5),
        legend.position = 'right',
        panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white")) +
  labs(title = title,
       subtitle = subtitle,
       x = x_title,
       y = y_title,
       shape="")+
  guides(color = guide_legend(override.aes = list(shape = c("średnia"=16))))  +
  scale_color_tableau()
  
  ggsave(paste0(title_to_save, "_box_", pomiar, ".png"), width = 8, height = 6, plot = plot, dpi = 300)
  
  
}
```


w grupach


Dane
```{r}
var = "irisyna_ng_ml" 

grupa = c("badana 1", "badana 2", "kontrolna")

pomiar = c(1, 2, 3, 4)

title = "Iryzyna"
title_to_save = "iryzyna"


x_title = "pomiary"
y_title = "ng/ml"
```

```{r warning=F}
for (i in grupa){
  
  
  grupy = c(i)
  subtitle = paste("grupa", grupy)

  plot <- df %>% 
      filter(
        id_pomiaru %in% pomiar,
        grupa %in% grupy) %>% 
    ggplot(aes(x = id_pomiaru, y = .data[[var]], color = id_pomiaru)) +
     geom_boxplot(alpha = 0.5, 
            show.legend = F) +
  stat_summary(fun.y=mean, geom="point", size=3, color = "red", aes(shape = "średnia"))+
  scale_shape_manual(values = c("średnia"=20))+  
  geom_sina(aes(color = id_pomiaru),
            alpha = 0.25,
            show.legend = F)+
    theme_classic() +
    theme(plot.title = element_text(face = "bold.italic", hjust = 0.5),
          plot.subtitle = element_text(face = 'italic', hjust = 0.5),
          legend.position = 'right',
          panel.background = element_rect(fill = "white"),
          plot.background = element_rect(fill = "white"),
          legend.background = element_rect(fill = "white")) +
    labs(title = title,
         subtitle = subtitle,
         x = x_title,
         y = y_title,
         shape = "")+
    guides(color = guide_legend(override.aes = list(shape = c("średnia"=16))))  +
    scale_color_tableau()
  
  ggsave(paste0(title_to_save, "_box_", grupy, ".png"), width = 8, height = 6, plot = plot, dpi = 300)

}
```
